---
title: "machine learning project"
author: "omar marey"
date: "3/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## load libraries 
```{r,echo=FALSE}
library(caret)
library(dplyr)

```
## downloading and loading the data 
```{r,results='hide'}
#download.file('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv','pml-training.csv')
#download.file('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv','pml-testing.csv')

training <- read.csv('pml-training.csv',na.strings = "#DIV/0!")
testing <- read.csv('pml-testing.csv',na.strings = "#DIV/0!")

```
it appears that the data has some null values coded as "#DIV/0!"
## exploring the data 
```{r}
summary(training)
str(training)
#unique(training)
```


## cleaning the data 
```{r,results='hide'}
cleaning <- function(Data,test = F){
  for(i in c(8:ncol(Data)-1)) {Data[,i] = as.numeric(as.character(Data[,i]))}
  complete_cols <- colnames(Data[colSums(is.na(Data)) == 0])[-(1:7)]
  Data <- Data[complete_cols]
  if(!test){
    Data$classe <- as.factor(Data$classe) }
  
  Data
}
training <- cleaning(training)
testing <- cleaning(testing,test = T)


```
here i made a function to clean the data it remove the first 2 columns as we will not need them in the 
model and convert the rest of the column to numerical values and remove any incomplete column (has any null values)

## model creation
```{r}
library(randomForest)
set.seed(47)
inTrain <- createDataPartition(training$classe, p=0.70, list=F)
trainData <- training[inTrain, ]
val <- training[-inTrain, ]
modfit <- randomForest(classe~.,data = trainData)

```
here i used random forest algorithm and divided the data to training set and validation set 

## validating the model
```{r}
pred <- predict(modfit,val)

confusionMatrix(val$classe,pred)
```

here we can see the result of prediction on the validation set and the model working pretty well

## quiz answers 
```{r}

res <- predict(modfit,testing)
res
```
those are the right answers of the quiz 
